---
title: "eda_utilsR_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{eda_utilsR_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, message=FALSE, warning=FALSE}
library(eda.utilsR)
```

As data rarely comes ready to be used and analyzed for machine learning right away, this package aims to help speed up the process of cleaning and doing initial exploratory data analysis (EDA). The package focuses on the tasks of dealing with outlier and missing values, scaling and correlation visualization.

This package contains four functions:
-   `cor_map`: A function to plot a correlation matrix of numeric
    columns in the dataframe
-   `outlier_identifier`: A function to identify and deal with outliers
-   `scale` A function to scale numerical values in the dataset
-   `imputer`: A function to impute missing values

### Data

For this demonstration we will use the following datasets.

```{r Dataset}
data <- data.frame('SepalLengthCm' = (c(5.1, 4.9, 4.7)), 
                   'SepalWidthCm'= (c(1.4, 1.4, 1.3)),
                  'PetalWidthCm'= (c(0.2, 0.1, 0.2)),
                  'Species' = (c('Iris-setosa','Iris-virginica', 'Iris-germanica')))

data_with_NA <- data.frame('SepalLengthCm' = (c(5.1, 4.9, 4.7)), 
                   'SepalWidthCm'= (c(1.4, 1.4, 1.3)),
                  'PetalWidthCm'= (c(0.2, 0.1, NA)))



data_with_outlier <- data.frame('SepalLengthCm' = (c(5.1, 4.9, 4.7, 5.2, 5.1, 5.2, 5.1, 4.8, 4.5)), 
                   'SepalWidthCm'= (c(1.4, 1.4, 1.3, 1.2, 1.2, 1.3, 1.6, 1.3, 1.2)),
                  'PetalWidthCm'= (c(0.2, 0.1, 30, 0.2, 0.3, 0.1, 0.4, 0.5, 0.3)))


```


### `imputer`

- __Impute__: Resolve skewed data by identifying missing data and outlier and provide corresponding remedy.


```{r imputer demo}
# calling imputer function
imputer(data_with_NA)
```


### `outlier_identifier`

- __Identify Outliers__: Identify and deal with outliers in the dataset.

Outlier can can affect our machine learning model quite a bit, especially on regression models since fitting the out scaled data point can dramatically increase our error. This function could help you by eliminate ourlier points or impute ourlier values with mean or median. So that the machine learning model can better catch the overall general trend of data.


```{r outlier_identifier demo}
# calling outlier_identifier function
outlier_identifier(data_with_outlier, method = "mean")
```



### `cor_map`

- __Correlation Heatmap Plotting__: Easily plot a correlation matrix along with its values to help explore data.

With EDA we generally want a quick snapshot of the data in order to understand it better. Visualizing the correlation between the numerical variables in the dataset are a good way to do that. This function allows you to quickly plot a correlation heat map of a chosen color scheme, along with correlation values overlaid. Users simply need to input the dataframe, the numeric columns wished to be plot () and an optional color scheme. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library('dplyr')
library('reshape2')
library('ggplot2')
```


```{r cor_map demo}
# defining the numeric columns
num_col <- c('SepalLengthCm', 'SepalWidthCm', 'PetalWidthCm')

# calling cor_map function
cor_map(data, num_col, "purpleorange")
```

### `scale`

- __Scaling__: Scale the data in preperation for future use in machine learning projects.


```{r scale demo}
# defining the numeric columns
num_col <- c('SepalLengthCm', 'SepalWidthCm', 'PetalWidthCm')

# calling scale function
scale(data, num_col, "minmax")
```

